---
title: "Analysis of salaries of data scientists"
author: "Group 12"
date: "December 1st, 2024"
output: pdf_document
---
-   Egbo Joseph (143037470)
-   Goudy Joshua (169031329)
-   Thambiaiah Melissa (169060509)
-   Umar Emaan (169108097)

```{r setup}
suppressWarnings(suppressPackageStartupMessages({
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(scatterplot3d)
library(corrr)
library(ggridges)
library(arrow)
library(tidymodels)
library(viridis)
library(ggrepel)
library(patchwork)}))
```
```{r import}
grapes_0<-read_csv('GRAPE_QUALITY.csv')
```
```{r data_cleaning}
grapes<-grapes_0|>
  select(!(sample_id))
grapes$quality_category<-factor(grapes$quality_category, levels = 
                                  c('Low', 'Medium', 'High', 'Premium'))
grapes<-grapes|> mutate(wine_type = case_when(
        variety == 'Riesling' ~ 'White',
        variety == 'Pinot Noir' ~ 'Red',
        variety == 'Sauvignon Blanc'~'White',
        variety == 'Merlot'~'Red',
        variety == 'Zinfandel'~'Red',
        variety == 'Chardonnay'~'White',
        variety == 'Syrah'~'Red',
        variety == 'Cabernet Sauvignon'~'Red',
        .default = variety),
      harvest_week = as.integer(week(harvest_date)) - 30 )
grapes<-grapes|> mutate(size_category = case_when(
        berry_size_mm < 15 ~ 'small',
        berry_size_mm < 20 ~ 'medium',
        berry_size_mm >= 20 ~ 'large',
        .default = 'uncategorized' ))
grapes$size_category<-factor(grapes$size_category, levels = c('small', 'medium', 'large'))
```
```{r training setup}
grapes_split <- initial_validation_split(grapes, prop = c(0.6, 0.2))
grapes_train <- training(grapes_split)
```
```{r, correlation table}
generate_correlation_table <- function(data) {
  continuous_vars <- data %>% select_if(is.numeric)
  correlation_matrix <- cor(continuous_vars, use = "complete.obs") 
  correlation_table <- as.data.frame(correlation_matrix) 
  return(correlation_table) }
grapes_woqs<-grapes_train|>select(!(quality_score))

round_table <- function(data, digits = 1) { data %>% mutate(across(everything(),
                                                                   ~ round(.x, digits))) }
corrs <- generate_correlation_table(grapes_woqs)
corrs <- round_table(corrs, 4)
rownames(corrs)<-c('Sugar Content', 'Acidity(PH)', 'Cluster Weight(g)', 'Berry Size(mm)',
                   'Sun Hours', 'Soil Moisture (%)', 'Rainfal(mm)', 'Harvest Week')
a<-kable(corrs[1:4], col.names=c('Sugar Content','Acidity(PH)', 'Cluster Weight(g)', 
                                 'Berry Size(mm)'))|> kable_styling()|>
  column_spec(1, width = '10em', background = rgb(0.878, 0.439, 0.49))|>
  column_spec(c(2,4), width = '10em', background = rgb(0.741, 0.949, 0.929))|>
  column_spec(c(3,5), width = '10em', background = rgb(1, 0.827, 0.827))|>
  row_spec(0, background = rgb(0.878, 0.439, 0.49))
b<-kable(corrs[4:7], col.names = c('Sun Hours','Soil Moisture(%)', 'Rainfal(mm)', 'harvest Week'
))|> kable_styling()|>
  column_spec(1, width = '10em', background = rgb(0.878, 0.439, 0.49))|>
  column_spec(c(2,4), width = '10em', background = rgb(0.741, 0.949, 0.929))|>
  column_spec(c(3,5), width = '10em', background = rgb(1, 0.827, 0.827))|>
  row_spec(0, background = rgb(0.878, 0.439, 0.49))
```
```{r, echo=FALSE}
a
b
```
# Exploratory Plot 1/Table 1 (only using 2 or more features, not the target variable)
```{r, sugar by region}
ggplot(grapes_train)+
  geom_density(mapping = aes(x = sugar_content_brix, fill = variety), alpha = 0.5, 
               show.legend = FALSE)+
  facet_wrap(~variety)+
  labs(y = 'Frequency',
       x = 'Sugar Content',
       title = 'Sugar Content by Variety',
       subtitle = '1 Unit of Sugar Content equal to 1g sugar per 100g of liquid',
       caption = 'Source: see footnote 1')
```

```{r, quality vs. sugar}
ggplot(grapes_train,mapping=aes(x = sugar_content_brix, y = quality_score))+
  geom_point(mapping = aes(colour = sun_exposure_hours))+
  geom_smooth(method = 'lm', formula = 'y~x', se = FALSE)+
  scale_color_viridis(option = "C") + 
  facet_wrap(~variety)+
  labs(x = 'Sugar Content',
       y = 'Quality Score',
       title = 'Quality vs. Sugar Content',
       subtitle = '1 Unit of Sugar Content equal to 1g sugar per 100g of liquid',
       caption = 'Source: see footnote 1',
       colour = 'Sun Exposure Hours')
```
# Model Plot 2: (only using 2 or more features, not the target variable)
```{r, violin plot}
ggplot(grapes_train, aes(y = size_category, x = quality_score))+
  geom_violin(aes(fill = size_category), show.legend=FALSE)+
  labs(title = 'Size Category Vs. Quality Score',
       subtitle = 'Small: 0-15mm, Medium: 15-20mm, Large: 20-25mm',
       x = 'Quality Score',
       y = 'Size Category',
       caption = 'Source: see footnote 1')
```
# Exploratory Linear Model 1: (only using 2 or more features, not the target variable)
```{r, grape_recipe}
recipe_1 <- recipe(quality_score ~ sugar_content_brix + size_category + wine_type,
                   data = grapes_train)|>step_dummy(wine_type)|>step_interact(~size_category)
recipe_2 <- recipe(quality_score ~ sugar_content_brix + size_category,
                   data = grapes_train) |>step_interact(~size_category)
recipe_3 <- recipe(quality_score ~ sugar_content_brix + wine_type,
                   data = grapes_train) |>step_dummy(wine_type) 
recipe_4 <- recipe(quality_score ~ size_category + wine_type,
                   data = grapes_train)|>step_dummy(wine_type)|>step_interact(~size_category)
grapes_recipe <- list(Sugar_Size_Type = recipe_1,Sugar_Size = recipe_2, 
                      Sugar_Type = recipe_3,Size_Type = recipe_4)
```
```{r grapes_workflow_set}
grapes_workflow_set <- workflow_set(grapes_recipe, models = list(lm = linear_reg()))
```
```{r, grapes_fit}
set.seed(1000)
grapes_fit <- workflow_set(
  grapes_recipe,
  models = list(lm = linear_reg())) |>
  workflow_map(fn = "tune_grid",resamples = validation_set(grapes_split)) 
grapes_fit |>
    collect_metrics()|>
    ggplot() +
    aes(x = wflow_id, y = mean) +
    geom_col(fill = "lightgrey", colour = "black") +
    facet_wrap(~ .metric, scales = "free")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, warning=FALSE}
train_test <- initial_split(grapes, prop = 0.8)
rf <- rand_forest(trees = 1000, mtry = tune(), min_n = 4) |>
    set_engine("randomForest") |>
    set_mode("regression")
tune_grid <- tibble(mtry = 1:4)
lm_versus_rf <- workflow_set(
 grapes_recipe,
  models = list(lm = linear_reg(), rf = rf)) |>
    workflow_map(fn = "tune_grid",grid = tune_grid,seed = 100,
        control = control_grid(save_pred = TRUE, save_workflow = TRUE),
        resamples = vfold_cv(training(grapes_split), v = 5))
lm_versus_rf |>
    autoplot(select_best = TRUE, metric = "rmse") +
    geom_label_repel(aes(label = wflow_id))
```
# Exploratory Linear Model 2: (only using 2 or more features, not the target variable)
```{r, grapes_recipe_2}
recipe_5 <- recipe(quality_score ~ sun_exposure_hours + region + variety,
                   data = grapes_train)|>step_interact(~region) |> step_interact(~variety)
recipe_6 <- recipe(quality_score ~ sun_exposure_hours + region,
                   data = grapes_train) |> step_interact(~region)
recipe_7 <- recipe(quality_score ~ sun_exposure_hours + variety,
                   data = grapes_train) |>step_interact(~variety) 
grapes_recipe_2 <- list(Sugar_variety_region = recipe_5,
                        Sugar_region = recipe_6, Sugar_variety = recipe_7)
```
```{r grapes_workflow_set_2}
grapes_workflow_set_2 <- workflow_set(grapes_recipe_2, models = list(lm = linear_reg()))
```
```{r, grapes_fit_2}
set.seed(1000)
grapes_fit_2 <- workflow_set(
  grapes_recipe_2,
  models = list(lm = linear_reg())) |>
  workflow_map(fn = "tune_grid",resamples = validation_set(grapes_split)) 
grapes_fit_2 |>
    collect_metrics()|>
    ggplot() +
    aes(x = wflow_id, y = mean) +
    geom_col(fill = "lightgrey", colour = "black") +
    facet_wrap(~ .metric, scales = "free")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, warning=FALSE}
train_test <- initial_split(grapes, prop = 0.8)
rf_2 <- rand_forest(trees = 1000, mtry = tune(), min_n = 3) |>
    set_engine("randomForest") |>
    set_mode("regression")
tune_grid_2 <- tibble(mtry = 1:3)
lm_versus_rf_2 <- workflow_set(
 grapes_recipe_2,
  models = list(lm = linear_reg(), rf = rf)) |>
    workflow_map(
        fn = "tune_grid",
        grid = tune_grid_2,
        seed = 100,
        control = control_grid(save_pred = TRUE, save_workflow = TRUE),
        resamples = vfold_cv(training(grapes_split), v = 5))
lm_versus_rf_2 |>
    autoplot(select_best = TRUE, metric = "rmse") +
    geom_label_repel(aes(label = wflow_id))
```
```{r best_model, error=TRUE}

best_lm <- lm_versus_rf |>extract_workflow("Sugar_Size_lm")
best_rf <- lm_versus_rf |>extract_workflow("Sugar_Size_rf")
```
```{r, test_fit}
sslm<-lm_versus_rf|>extract_workflow_set_result('Sugar_Size_lm')|>
  select_best(metric='rmse')
ssrf<-lm_versus_rf|>extract_workflow_set_result('Sugar_Size_rf')|>
  select_best(metric='rmse')
test_lm<-lm_versus_rf|>
  extract_workflow('Sugar_Size_lm')|>
  finalize_workflow(sslm)|>
  last_fit(split = grapes_split)
test_rf<-lm_versus_rf|>
  extract_workflow('Sugar_Size_rf')|>
  finalize_workflow(ssrf)|>
  last_fit(split = grapes_split)
point<-bind_rows(
    lm = collect_predictions(test_lm),
    rf = collect_predictions(test_rf),
    .id = "model") |>
    ggplot() +
    aes(x = quality_score, y = .pred, colour = model) +
    geom_point(shape = 1)+
  labs(title = 'Quality Score Model Predictions',
    y = 'Predicted Quality',
    x = 'Actual Quality')
bar<-bind_rows(
    lm = test_lm,
    rf = test_rf,
    .id = "model") |>
    unnest_wider(.metrics) |>
    unnest_longer(c(.metric, .estimator, .estimate, .config)) |>
    ggplot() +
    aes(x = model, y = .estimate) +
    geom_col(fill = "lightgrey", color = 1) +
    facet_wrap(~ .metric, scales = "free_y")+
    labs(title = 'RMSE and RSQ of Models',
         y = 'Estimate',
         x = 'Model')
bar + point
```
^[Vishal. (2024). Grape Quality [Data set]. Kaggle. https://doi.org/10.34740/KAGGLE/DSV/9781093]
